Calendar-based graphics for visualizing people's daily schedules

### 2019-05-16 main.Rmd
new cover letter for resubmission @dicook can you pls take a look at the cover letter?
revised section 2.3

## Variations {#sec:variations}

### Overlaying and faceting subsets

Plots can be layered. The comparison of sensors can be done by overlaying the values for each (Figure \@ref(fig:overlay)). Differences between the pedestrian patterns at these sensors can be seen. Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays. This suggests that Flagstaff Station has limited functionality on non-work days. From Figure \@ref(fig:overlay) it can be seen that Birrarung Marr has a distinct temporal pattern from the other two all year round. The nighttime events, such as White Night, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.

@@

(ref:overlay-cap) Overlaying line graphs of the three sensors in the monthly calendar. The three sensors demonstrate very different traffic patterns. Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station.

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \@ref(fig:facet) presents the idea of faceting calendar plots by sensors. This allows comparing the overall structure between sensors, while emphasizing individual sensor variation. In particular, it can be immediately learned that Birrarung Marr was busy and packed over many weekends, but events took place on Friday evenings only in September. The Australian Open, a major international tennis tournament, attracted much traffic constantly in the last two weeks of January.

@@

(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized, and the shapes can be compared across the cells and sensors.

@@

The `frame_calendar()` function is not constrained to line plots. The full range of plotting capabilities in **ggplot2** is essentially available. Figure \@ref(fig:scatterplot) shows lag scatterplots with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. This figure is organized in the daily calendar layout. Figure \@ref(fig:scatterplot) indicates strong autocorrelation on weekends, and weak autocorrelation on work days. At the higher counts on work days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs. This is due to the morning and afternoon commuter patterns in Figure \@ref(fig:fs-2016): the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.

@@

# Making calendar facets {#sec:facet-calendar}

@@

# Acknowledgments {-}

>>>>

## Varieties of calendar display {#sec:variations}

### Information overlay

Plots can be layered. A comparison of sensors can be done by overlaying them in the same calendar pane. Figure \@ref(fig:overlay)) overlays the pedestrian counts for three locations on the same calendar. Differences between the pedestrian patterns at these locations can be more directly compared. For example, the magnitude of the difference in pedestrians at Flagstaff Station at peak hours of commuter can be seen. The big peak in pedestrian counts for special events at Birrarung Marr is clear. Birrarung Marr has a very distinct temporal pattern relative to the other two locations. The nighttime events, such as White Night (third Saturday in February), only affects the foot traffic at the State Library and Birrarung Marr.

@@

(ref:overlay-cap) Overlaying line graphs of the three sensors in the monthly calendar, to enable a direct comparison of the counts at three locations. Theey have very different traffic patterns. Birrarung Marr tends to attract large numbers of pedestrians for special events typically held on weekends, contrasting to the bimodal massive peaks shoing commuting traffic at Flagstaff Station.

### Faceting by covariates

To avoid overlapping, when differences between groups are large enough to be seen separately, the calendar layout can be facetted into a series of subplots for the different sensors. Figure \@ref(fig:facet) shows calendar plots that are facetted by sensors. This arrangement allows comparison of the overall structure between sensors, while emphasizing individual sensor variation. In particular, it can be immediately learned that Birrarung Marr was busy and packed over many weekends, but events took place on Friday evenings only in September. The Australian Open, a major international tennis tournament, attracted constant foot traffic in the last two weeks of January. The calendar plot can be facetted by any categorical variable in the data.

@@

(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasized.

@@

Almost any type of plot can be shown in a calendar pane. Most of the full range of plotting capabilities in **ggplot2** is available. An example is shown in Figure \@ref(fig:scatterplot): the panes contain lag scatterplots, constructed with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. It indicates strong autocorrelation on weekends, and weak autocorrelation on work days. The the V-shape in the week day graphs, arises when the next hour sees a substantial increase or decrease in counts. This is due to the morning and afternoon commuter patterns in Figure \@ref(fig:fs-2016): the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.

@@

### Facetted calendar {#sec:facet-calendar}

@@

# Acknowledgements {-}
